Transformation of forecasts for evaluating predictive performance in an epidemiological context

### 2023-06-26 second submission after review/reply_to_reviews.tex

\indented
A natural question is which other transformations could be applied and whether resulting scores remain (strictly) proper. In principle, any transformation function can be applied simultaneously to forecasts and observations as long as the definition of the transformation is independent of the forecasts and any quantities unknown at the time of forecasting, including the observed value. This simply corresponds to a re-definition of the forecasting target. However, applying non-invertible transformations leads to a loss in information conveyed by forecasts, which we consider undesirable. The resulting score will be proper, but it may not be strictly proper anymore (as forecasts differing from the forecaster's true belief on the original scale may be identical on the transformed scale). When using the CRPS or the WIS, it seems most appropriate to use only strictly monotonic transformations such as the natural logarithm or the square root as otherwise the encoded notion of distance may become meaningless.

Some other strictly monotonic transformations that can be applied are scaling by the population size or scaling by past observations. The latter, as discussed in Section 4, is similar to applying a log-transformation, but corresponds to evaluating a forecast of multiplicative, rather than exponential growth rates. The arising issue of dividing by zero can again be solved by adding a small offset $a$. Scaling a forecast by the later observed value (as opposed to scaling by past observations) is generally not permissible as it can result in improper scores (see \citealt{lerchForecasterDilemmaExtreme2015} on the closely related topic of weighting scores with a function of the observed value). Similarly, scaling forecasts and observations by a function of the predictive distribution (like the predictive mean) may lead to improper scores; however, we are unaware of existing theoretical arguments on this.

When applying a transformation, the order of the operations matters, and applying a transformation after scores have been computed generally does not guarantee that the score remains proper. In the case of log transforms, taking the logarithm of the CRPS values, rather than scoring the log-transformed forecasts and data, results in an improper score. We illustrate this point using simulated data in Figure SI.1, where it can be seen that in the example overconfident models perform best in terms of the log WIS. We note that strictly speaking, re-scaling average scores by the average score of a baseline model or average scores across different models  to obtain skill scores likewise leads to improper scores \citep{gneitingStrictlyProperScoring2007}. The application of such skill scores, however, is established practice and considered largely unproblematic.

@@

\notindented

\blue
Figure 4: The data (not) selected seems to obscure the key period of predicting the start of a case wave, as we expect the CRPS of log data to emphasize missed case increase predictions there more heavily than the natural scale. Deaths are not a substitute because they have access to a leading indicator and the same sort of misses may not be observed. The evaluations here may also miss some overshooting of the case peak. Which rules from Table SI.2 are coming into play and why? Is there a more complete alternative example?

\black
The data that was removed here was largely due to data revisions, not due to forecasts removed by the filtering on our part. 
We added the percentages of observed values in the main text and added the following two Figures in the SI to make that clearer.

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.49\textwidth]{../output/figures/number-anomalies.png}
     \includegraphics[width=0.49\textwidth]{../output/figures/erroneous-forecasts.png}
     \caption{}
    \label{fig:number-anomalies}
\end{figure}

@@

9. Table SI.2, "Criteria for removing forecasts": forecasts were removed if their median prediction differed greatly from the true value. This should probably be mentioned in the manuscript text, somewhere around line 235 (where Table SI.2 is referenced), to notify the reader that the definition of "erroneous forecasts" includes forecasts that are in extremely poor agreement with the ground truth. Otherwise, I feel that "erroneous" may be open to misinterpretation (e.g., only removing forecasts that predicted negative counts, NaN values, etc).

\black
This is a good point. We changed the sentence to

In addition, we filtered out a small number of erroneous forecasts that were in extremely poor agreement with the observed data, as defined by any of the conditions listed in Table SI.2.

\clearpage
\bibliography{../bib/log-or-not-v2.bib, ../bib/log-or-not.bib, ../bib/software.bib}

### 2023-06-26 first submission/Manuscript_first_submission.tex
Prepare files for second submission

\title{Scoring epidemiological forecasts on transformed scales}
  \author{Nikos I. Bosse\thanks{Department of Infectious Disease Epidemiology, London School of Hygiene \& Tropical Medicine, London, United Kingdom} $^{,}$\thanks{Centre for the Mathematical Modelling of Infectious Diseases, London, United Kingdom} $^{,}$\thanks{NIHR Health Protection Research Unit in Modelling \& Health Economics} $^{,*}$,  
  Sam Abbott\footnotemark[1] $^{,}$\footnotemark[2]$ ^{}$, 
  Anne Cori\thanks{MRC Centre for Outbreak Analysis and Modelling, Department of Infectious Disease Epidemiology, School of Public Health, Imperial College London, London, United Kingdom} $^{}$, \\
  Edwin van Leeuwen\footnotemark[1] $^{,}$\thanks{Modelling \& Economics Unit and NIHR Health Protection Research Unit in Modelling \& Health Economics, UK Health Security Agency, London, United Kingdom} $^{}$, 
  Johannes Bracher\thanks{Chair of Statistical Methods and Econometrics, Karlsruhe Institute of Technology, Karlsruhe, Germany } $^{,}$\thanks{Computational Statistics Group, Heidelberg Institute for Theoretical Studies, Heidelberg, Germany } $^{, \dagger}$, 
  Sebastian Funk\footnotemark[1] $^{, }$\footnotemark[2] $^{, }$\footnotemark[3] $ ^{, \dagger}$}

@@

Forecast evaluation is essential for the development of predictive epidemic models and can inform their use for public health decision-making. Common scores to evaluate epidemiological forecasts are the Continuous Ranked Probability Score (CRPS) and the Weighted Interval Score (WIS), which can be seen as measures of the absolute distance between the forecast distribution and the observation. However, applying these scores directly to predicted and observed incidence counts may not be the most appropriate due to the exponential nature of epidemic processes and the varying magnitudes of observed values across space and time. In this paper, we argue that transforming counts before applying scores such as the CRPS or WIS can effectively mitigate these difficulties and yield epidemiologically meaningful and easily interpretable results. Using the CRPS on log-transformed values as an example, we list three attractive properties: Firstly, it can be interpreted as a probabilistic version of a relative error. Secondly, it reflects how well models predicted the time-varying epidemic growth rate. And lastly, using arguments on variance-stabilizing transformations, it can be shown that under the assumption of a quadratic mean-variance relationship, the logarithmic transformation leads to expected CRPS values which are independent of the order of magnitude of the predicted quantity. Applying a transformation of log(x + 1) to data and forecasts from the European COVID-19 Forecast Hub, we find that it changes model rankings regardless of stratification by forecast date, location or target types. Situations in which models missed the beginning of upward swings are more strongly emphasised while failing to predict a downturn following a peak is less severely penalised when scoring transformed forecasts as opposed to untransformed ones. We conclude that appropriate transformations, of which the natural logarithm is only one particularly attractive option, should be considered when assessing the performance of different models in the context of infectious disease incidence.
\end{abstract}

@@

Probabilistic forecasts \citep{heldProbabilisticForecastingInfectious2017} play an important role in decision-making in epidemiology and public health \citep{doi:10.2105/AJPH.2022.306831}, as well as other areas as diverse as economics \citep{timmermannForecastingMethodsFinance2018} or meteorology \citep{gneitingWeatherForecastingEnsemble2005}. Forecasts based on epidemiological modelling in particular have received widespread attention during the COVID-19 pandemic. Evaluations of forecasts can provide feedback for researchers to improve their models and train ensembles. They moreover help decision-makers distinguish good from bad predictions and choose forecasters and models that are best suited to inform future decisions.

Probabilistic forecasts are usually evaluated using so-called (strictly) proper scoring rules \citep{gneitingStrictlyProperScoring2007}, which return a numerical score as a function of the forecast and the observed data. 
Proper scoring rules are constructed such that they encourage honest forecasting and cannot be `gamed' or `cheated'. 
Assuming that the forecaster's actual best judgement corresponds to a predictive distribution $F$, a proper score is constructed such that if $F$ was the data-generating process, no other distribution $G$ would yield a better expected score. A scoring rule is called \textit{strictly} proper if there is no other distribution that under $F$ achieves the \textit{same} expected score as $F$, meaning that any deviation from $F$ leads to a worsening of expected scores. Forecasters (anyone or anything that issues a forecast) are thus incentivised to report their true belief $F$ about the future. 
Common proper scoring rules are the logarithmic or log score \citep{goodRationalDecisions1952} and the continuous ranked probability score~\citep[CRPS,][]{gneitingStrictlyProperScoring2007}. The log score is the predictive log density or probability mass evaluated at the observed value. It is supported by the likelihood principle \citep{Winkler1996} and has many desirable theoretical properties; however, the particularly severe penalties it assigns to occasional misguided forecasts make it little robust \citep{bracherEvaluatingEpidemicForecasts2021}. Moreover, it is not easily applied to forecasts reported as samples or quantiles, as used in many recent disease forecasting efforts. It is nonetheless occasionally used in epidemiology (see e.g., \citealt{heldProbabilisticForecastingInfectious2017, Johansson2019}), but in recent years the CRPS and the weighted interval score ~\citep[WIS,][]{bracherEvaluatingEpidemicForecasts2021} have become increasingly popular.

The CRPS measures the distance of the predictive distribution to the observed data as 
\begin{linenomath*}
\begin{equation}
    \text{CRPS}(F, y) = \int_{-\infty}^\infty \left( F(x) - \boldsymbol{1}(x \geq y) \right)^2 dx,
\end{equation}    
\end{linenomath*}
where $y$ is the true observed value, $F$ is the cumulative distribution function (CDF) of the predictive distribution, and $\boldsymbol{1}()$ is the indicator function. The CRPS can be understood as a generalisation of the absolute error to predictive distributions, and interpreted on the natural scale of the data. The WIS is an approximation of the CRPS for predictive distributions represented by a set of predictive quantiles and is currently used to assess forecasts in the so-called COVID-19 Forecast Hubs in the US \citep{cramerCOVID19ForecastHub2020, cramerEvaluationIndividualEnsemble2021}, Europe \citep{sherrattPredictivePerformanceMultimodel2022} and Germany and Poland \citep{bracherShorttermForecastingCOVID192021, bracherNationalSubnationalShortterm2021}, as well as the US \textit{FluSight} project on influenza forecasting\citep{CdcepiFlusightforecastdata2022}. The WIS is defined as 
\begin{linenomath*}
\begin{equation}
    \text{WIS}(F, y) = \frac{1}{K} \times \sum_{k = 1}^{K} 2 \times \left[ \boldsymbol{1}(y \leq q_{\tau_k}) - \tau_k \right] \times ( q_{\tau_k} - y), 
\end{equation}
\end{linenomath*}
where $q_{\tau}$ is the $\tau$ quantile of the forecast $F$, $y$ is the observed outcome and $K$ is the number of (roughly equally spaced) predictive quantiles provided. The WIS can be decomposed into three components, dispersion, underprediction and overprediction, which reflect the spread of the forecast and whether it was centred above or below the observed value. We show an alternative definition based on central prediction intervals in Supplement \ref{sec:alternative-wis} which illustrates this decomposition.

The notion of absolute distance encoded by the CRPS and WIS provides a straightforward interpretation, but may not always be the most useful perspective in the context of infectious disease spread. Especially in their early phase, outbreaks are best conceived as exponential processes, characterized by potentially time varying reproduction numbers $R_t$ \citep{gosticPracticalConsiderationsMeasuring2020} or epidemic growth rates $r_t$ \citep{dushoffSpeedStrengthEpidemic2021}. If the true modelling task revolves around estimating and forecasting these quantities, then evaluating forecasts based on the absolute distance between forecasted and observed incidence values penalises underprediction (of the reproduction number or growth rate) less than overprediction by the same amount. For illustration, consider an incidence forecast issued at time 0 and referring to time $t$ that misses the correct average growth rate $\bar{r}_t$ by either $-\epsilon$ or $+\epsilon$. Then the ratio of the resulting absolute errors on the scale of observed incidences $y_{t}$ is

\begin{linenomath*}
\begin{equation}
\frac{\left|y_0 \exp[(\bar{r}_t - \epsilon) \times t] - y_0 \exp(\bar{r}_t t)\right|}{\left| y_0 \exp[(\bar{r}_t + \epsilon) \times t] - y_0 \exp(\bar{r}_t t) \right|} = \exp(-\epsilon t) < 1.
\end{equation}
\end{linenomath*}

% For exponential processes errors on the observed value grow exponentially with the error on the estimated reproduction number or growth rate. 
If one is to measure the ability to forecast the underlying infection dynamics, it may thus be more desirable to evaluate errors on the scale of the growth rate directly.

Another argument against using notions of absolute distance between predicted and observed incidence values is that forecast consumers may find errors on a relative scale easier to interpret and more useful in order to track predictive performance across targets of different orders of magnitude. 
\cite{bolinLocalScaleInvariance2021} have proposed the scaled CRPS (SCRPS) which is locally scale invariant; however, it does not correspond to a relative error measure and lacks a straightforward interpretation as available for the CRPS.

% A closely related aspect to relative scores (as opposed to absolute scores) is that in the evaluation one may wish to give similar weight to all 
Lastly, it may be considered desirable to give all forecast targets similar weight in an overall performance evaluation. As the CRPS typically scales with the order of magnitude of the quantity to be predicted, this is not the case for the CRPS, which will typically assign higher scores to forecast targets with high expected values (e.g., in large locations or around the peak of an epidemic). \cite{bracherEvaluatingEpidemicForecasts2021} have argued that this is a desirable feature, directing attention to situations of particular public health relevance. An evaluation based on absolute errors, however, will assign little weight to other potentially important aspects, such as the ability to correctly predict future upswings while observed numbers are still low.

In many fields, it is common practice to forecast transformed quantities (see e.g. \cite{taylorEvaluatingVolatilityInterval1999} in finance, \cite{mayrLogLevelVAR2015} in macroeconomics, \cite{loweStochasticRainfallrunoffForecasting2014} in hydrology or \cite{fuglstadDoesNonstationarySpatial2015} in meteorology). While the goal of the transformations is often to improve the accuracy of the predictions, they can also be used to enhance and complement the evaluation process. 
In this paper, we argue that the aforementioned issues with evaluating epidemic forecasts based on measures of absolute error on the natural scale can be addressed by transforming the forecasts and observations prior to scoring using some strictly monotonic transformation. Strictly monotonic transformations can shift the focus of the evaluation in a way that may be more appropriate for epidemiological forecasts, while guaranteeing that the score remains proper. Many different transformations may be appropriate and useful, depending on the exact context, the desired focus of the evaluation, and specific aspects forecast consumers care most about (see a broader discussion in Section \ref{sec:discussion}).

For conceptual clarity and to allow for a more in-depth discussion, we focus mostly on the natural logarithm as a particularly attractive transformation in the context of epidemic phenomena. We refer to this transformation as 'log-transformation' and to scores that have been computed from log-transformed forecasts and observations as scores 'on the log scale' (as opposed to scores 'on the natural scale', which involve no transformation). In the theoretical discussion in Section \ref{sec:methods}, 'log-transformation' and 'log scale' generally refer to a transformation of $\log_{e}(x)$. For practical applications (Section \ref{sec:HUB}) we also use these terms to describe a transformation of $\log_{e}(x + a)$ with a small $a > 0$ in order to keep the terminology and notation simple. For a prediction target with strictly positive support, the CRPS after applying a log-transformation is given by
%
\begin{linenomath*}
\begin{equation}
    \text{CRPS}(F_{\log}, \log y) = \int_{-\infty}^\infty \left( F_{\log}(x) - \boldsymbol{1}(x \geq \log y) \right)^2 dx.
\end{equation}    
\end{linenomath*}
%
Here, $y$ is again the observed outcome and $F_{\log}$ is the predictive CDF of the log-transformed outcome, i.e.,
\begin{linenomath*}
\begin{equation}
F_{\log}(x) = F(\exp(x)),
\end{equation}    
\end{linenomath*}
with $F$ the CDF on the original scale. Instead of a score representing the magnitude of absolute errors, applying a log-transformation prior to the CRPS yields a score which a) measures relative error (see Section \ref{sec:methods:relative}), b) provides a measure for how well a forecast captures the exponential growth rate of the target quantity (see Section \ref{sec:methods:growthrate}) and c) is less dependent on the expected order of magnitude of the quantity to be predicted (see Section \ref{sec:methods:vst}). 
We therefore argue that such evaluations on the logarithmic scale should complement the prevailing evaluations on the natural scale. 
Other transformations may likewise be of interest. We briefly explore the square root transformation as an alternative transformation. 
Our analysis mostly focuses on the CRPS (or WIS) as an evaluation metric for probabilistic forecasts, given its widespread use throughout the COVID-19 pandemic. We note that the logarithmic score has scale invariance properties which imply that score differences between different forecasts are invariant to strictly monotonic transformations (see \citealt{Lehmann1950} on corresponding properties of likelihood ratios and \citealt{diksLikelihoodbasedScoringRules2011}). The question of the right scale to evaluate forecasts on does therefore not arise for the log score.

The remainder of the article is structured as follows. In Sections \ref{sec:methods:relative}--\ref{sec:methods:vst} we provide some mathematical intuition on applying the log-transformation prior to evaluating the CRPS, highlighting the connections to relative error measures, the epidemic growth rate and variance stabilizing transformations.
We then discuss the effect of the log-transformation on forecast rankings (Section \ref{sec:methods:rankings}) as well as practical considerations for applying transformations in general and the log-transformation in particular (Section \ref{sec:methods:considerations}). To analyse the real-world implications of the log-transformation we use forecasts submitted to the European COVID-19 Forecast Hub  \citep[ Section \ref{sec:HUB}]{europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022}. Finally, we provide scoring recommendations, discuss alternative transformations that may be useful in different contexts, and suggest further research avenues (Section \ref{sec:discussion}).

@@

To illustrate the effect of applying the natural logarithm prior to evaluating forecasts we consider the absolute error, which the CRPS and WIS generalize to probabilistic forecasts. We assume strictly positive support (meaning that no specific handling of zero values is needed), a restriction we will address when applying this transformation in practice. When considering a point forecast $\hat{y}$ for a quantity of interest $y$, such that 
%
\begin{linenomath*}
\begin{equation}
y = \hat{y} + \varepsilon,
\end{equation}
\end{linenomath*}
the absolute error is given by $|\varepsilon|$. When taking the logarithm of the forecast and the observation first, thus considering 
\begin{linenomath*}
\begin{equation}
\log y = \log \hat{y} + \varepsilon^*,
\end{equation}
\end{linenomath*}
the resulting absolute error $\left|\varepsilon^*\right|$ can be interpreted as an approximation of various common relative error measures. Using that $\log(a) \approx a - 1$ if $a$ is close to 1, we get
\begin{linenomath*}
\begin{equation}
|\varepsilon^*| = |\log \hat{y} - \log y| = \left|\log\left(\frac{\hat{y}}{y}\right) \right| \ \ \stackrel{\text{if } \hat{y} \ \approx \ y}{\approx} \ \ \left| \frac{\hat{y}}{y} - 1 \right| \ \ = \ \ \left| \frac{\hat{y} - y}{y} \right|.
\end{equation}
\end{linenomath*}
The absolute error after log transforming is thus an approximation of the \textit{absolute percentage error} \citep[APE,][]{gneitingMakingEvaluatingPoint2011a} as long as forecast and observation are close. As we assumed that $\hat{y} \approx y$, we can also interpret it as an approximation of the \textit{relative error} \citep[RE,][]{gneitingMakingEvaluatingPoint2011a}
\begin{linenomath*}
\begin{equation}
\left| \frac{\hat{y} - y}{\hat{y}} \right|
\end{equation}
\end{linenomath*}
and the \textit{symmetric absolute percentage error} (SAPE; see e.g., \citealt{Flores1986})
\begin{linenomath*}
\begin{equation}
\left| \ \frac{\hat{y} - y}{y/2 + \hat{y}/2} \ \right|.
\end{equation}
\end{linenomath*}
As Figure \ref{fig:SAPE} shows, the alignment with the SAPE is in fact the closest and holds quite well even if predicted and observed value differ by a factor of two or three. Generalising to probabilistic forecasts, the CRPS applied to log-transformed forecasts and outcomes can thus be seen as a probabilistic counterpart to the symmetric absolute percentage error, which offers an appealing intuitive interpretation.

@@

Another interpretation for the log-transform is possible if the generative process is framed as exponential with a time-varying growth rate $r(t)$~\citep[see, e.g.,][]{wallingaHowGenerationIntervals2007}, i.e.
\begin{linenomath*}
\begin{equation}
\frac{d}{dt}y(t) = r(t)y(t)
\end{equation}
\end{linenomath*}
%
which is solved by
%
\begin{linenomath*}
\begin{equation}
y(t) = y_0 \exp \left( \int_0^t r(t') dt' \right) = y_0 \exp (\bar{r_t}t)
\end{equation}
\end{linenomath*}
where $y_0$ is an initial data point and $\bar{r_t}$ is the mean of the growth rate between the initial time point $0$ and time $t$.

>>>>

\title{Transformation of forecasts for evaluating predictive performance in an epidemiological context}
  \author{Nikos I. Bosse\thanks{Department of Infectious Disease Epidemiology, London School of Hygiene \& Tropical Medicine, London, United Kingdom} $^{,}$\thanks{Centre for the Mathematical Modelling of Infectious Diseases, London, United Kingdom} $^{,}$\thanks{NIHR Health Protection Research Unit in Modelling \& Health Economics} $^{,*}$,  
  Sam Abbott\footnotemark[1] $^{,}$\footnotemark[2]$ ^{}$, 
  Anne Cori\thanks{MRC Centre for Outbreak Analysis and Modelling, Department of Infectious Disease Epidemiology, School of Public Health, Imperial College London, London, United Kingdom} $^{}$, \\
  Edwin van Leeuwen\footnotemark[1] $^{,}$\thanks{Modelling \& Economics Unit and NIHR Health Protection Research Unit in Modelling \& Health Economics, UK Health Security Agency, London, United Kingdom} $^{}$, 
  Johannes Bracher\thanks{Chair of Statistical Methods and Econometrics, Karlsruhe Institute of Technology, Karlsruhe, Germany } $^{,}$\thanks{Computational Statistics Group, Heidelberg Institute for Theoretical Studies, Heidelberg, Germany } $^{, \dagger}$, 
  Sebastian Funk\footnotemark[1] $^{, }$\footnotemark[2] $^{, }$\footnotemark[3] $ ^{, \dagger}$}

@@

Forecast evaluation plays an essential role in the development cycle of predictive epidemic models and can inform their use for public health decision-making. Common scores to evaluate epidemiological forecasts are the Continuous Ranked Probability Score (CRPS) and the Weighted Interval Score (WIS), which are both measures of the absolute distance between the forecast distribution and the observation. They are commonly applied directly to predicted and observed incidence counts, but it can be questioned whether this is the optimal procedure for comparing models given the exponential nature of epidemic processes and the several orders of magnitude that observed values can span over space and time. In this paper, we argue that transforming counts before applying scores such as the CRPS or WIS can effectively mitigate these difficulties and yield epidemiologically meaningful and easily interpretable results. We motivate the procedure threefold using the CRPS on log-transformed counts as an example: Firstly, it can be interpreted as a probabilistic version of a relative error. Secondly, it reflects how well models predicted the time-varying epidemic growth rate. And lastly, using arguments on variance-stabilizing transformations, it can be shown that under the assumption of a quadratic mean-variance relationship, the logarithmic transformation leads to expected CRPS values which are independent of the order of magnitude of the predicted quantity. Applying the log transformation to data and forecasts from the European COVID-19 Forecast Hub, we find that it changes model rankings regardless of stratification by forecast date, location or target types. Situations in which models missed the beginning of upward swings are more strongly emphasized while failing to predict a downturn following a peak is less severely penalized. We conclude that appropriate transformations, of which the natural logarithm is only one particularly attractive option, should be considered when assessing the performance of different models in the context of infectious disease incidence. 
\end{abstract}

@@

Probabilistic forecasts \citep{heldProbabilisticForecastingInfectious2017} play an important role in decision-making in epidemiology and public health \citep{doi:10.2105/AJPH.2022.306831}, as well as other areas as diverse as economics \citep{timmermannForecastingMethodsFinance2018} or meteorology \citep{gneitingWeatherForecastingEnsemble2005}. Forecasts based on epidemiological modelling in particular has received widespread attention during the COVID-19 pandemic. Evaluations of forecasts can provide feedback for researchers to improve their models and train ensembles. They moreover help decision-makers distinguish good from bad predictions and choose forecasters and models that are best suited to inform future decisions.

Probabilistic forecasts are usually evaluated using so-called proper scoring rules \citep{gneitingStrictlyProperScoring2007}, which return a numerical score as a function of the forecast and the observed data. Proper scoring rules are constructed such that forecasters (anyone or anything that issues a forecast) are incentivised to report their true belief about the future. Examples of proper scoring rules that have been used to assess epidemiological forecasts are the Continuous Ranked Probability Score~\citep[CRPS,][]{gneitingStrictlyProperScoring2007} or its discrete equivalent, the Ranked Probability Score~\citep[RPS,][]{funkAssessingPerformanceRealtime2019}, and the Weighted Interval Score \citep{bracherEvaluatingEpidemicForecasts2021}. 
The CRPS measures the distance of the predictive distribution to the observed data as 
\begin{linenomath*}
\begin{equation*}
    \text{CRPS}(F, y) = \int_{-\infty}^\infty \left( F(x) - 1(x \geq y) \right)^2 dx,
\end{equation*}    
\end{linenomath*}
where $y$ is the true observed value and $F$ the cumulative distribution function (CDF) of the predictive distribution. The CRPS can be understood as a generalisation of the absolute error to predictive distributions, and interpreted on the natural scale of the data. The WIS is an approximation of the CRPS for predictive distributions represented by a set of predictive quantiles and is currently used to assess forecasts in so-called COVID-19 Forecast Hubs in the US \citep{cramerCOVID19ForecastHub2020, cramerEvaluationIndividualEnsemble2021}, Europe \citep{sherrattPredictivePerformanceMultimodel2022}, Germany and Poland \citep{bracherShorttermForecastingCOVID192021, bracherNationalSubnationalShortterm2021}, as well as the US Influenza Forecasting Hub \citep{CdcepiFlusightforecastdata2022}. The WIS is defined as 
\begin{linenomath*}
\begin{equation*}
    \text{WIS}(F, y) = \frac{1}{K} \times \sum_{k = 1}^{K} 2 \times \left[ \boldsymbol{1}(y \leq q_{\tau_k}) - \tau_k \right] \times ( q_{\tau_k} - y), 
\end{equation*}
\end{linenomath*}
where $q_{\tau}$ is the $\tau$ quantile of the forecast $F$, $y$ is the observed outcome, $K$ is the number of predictive quantiles provided and $\boldsymbol{1}$ is the indicator function. The WIS can be decomposed into three components, dispersion, overprediction, underprediction, which reflect the width of the forecast and whether it was centred above or below the observed value. We show an alternative definition based on central prediction intervals in Section \ref{sec:alternative-wis} which illustrates this decomposition.

The dynamics of infectious processes are often described by the complementary concepts of the reproduction number $R$ \citep{gosticPracticalConsiderationsMeasuring2020} and growth rate $r$ \citep{wallingaHowGenerationIntervals2007}, where $R$ describes the strength and $r$ the speed of epidemic growth \citep{dushoffSpeedStrengthEpidemic2021}.
In the absence of changes in immunity, behaviour or other factors that may affect the intensity of transmission, the reproduction number would be expected to remain approximately constant.
In that case, the number of new infections in the population grows exponentially in time.
This behaviour was observed, for example, early in the COVID-19 pandemic in many countries~\citep{pellisChallengesControlCOVID192021}.

If case numbers are evolving based on an exponential process and the  modelling task revolves around estimating and forecasting the reproduction number or the corresponding growth rate, then evaluating forecasts based on the absolute distance between forecast and observed value penalises underprediction (of the reproduction number or growth rate) less than overprediction by the same amount. This is because for exponential processes errors on the observed value grow exponentially with the error on the estimated reproduction number or growth rate. 
If one is to measure the ability of forecasts to assess and forecast the underlying infection dynamics, it may thus be more desirable to evaluate errors on the growth rate directly.

Evaluating forecasts using the CRPS or WIS means that scores represent a measure of absolute errors. However, forecast consumers may find errors on a relative scale easier to interpret and more useful in order to track predictive performance across targets of different orders of magnitude. 
\cite{bolinLocalScaleInvariance2021} have proposed the scaled CRPS (SCRPS) which is locally scale invariant; however, it does not correspond to a relative error measure and lacks a straightforward interpretation as available for the CRPS.

A closely related aspect to relative scores (as opposed to absolute scores) is that in the evaluation one may wish to give similar weight to all considered forecast targets. As the CRPS typically scales with the order of magnitude of the quantity to be predicted, this is not the case for the CRPS, which will typically assign higher scores to forecast targets with high expected values (e.g., in large locations or around the peak of an epidemic). \cite{bracherEvaluatingEpidemicForecasts2021} have argued that this is a desirable feature, directing attention to situations of particular public health relevance. An evaluation based on absolute errors, however, will assign little weight to other potentially important aspects, such as the ability to correctly predict future upswings while observed numbers are still low.

In many fields, it is common practice to forecast transformed quantities (see e.g. \cite{taylorEvaluatingVolatilityInterval1999} in finance, \cite{mayrLogLevelVAR2015} in macroeconomics, \cite{loweStochasticRainfallrunoffForecasting2014} in hydrology or \cite{fuglstadDoesNonstationarySpatial2015} in meteorology). While the goal of the transformations is usually to improve the accuracy of the predictions, they can also be used to enhance and complement the evaluation process. 
In this paper, we argue that the aforementioned issues with evaluating epidemic forecasts based on measures of absolute error on the natural scale can be addressed by transforming the forecasts and observations prior to scoring using some strictly monotonic transformation. Strictly monotonic transformations can shift the focus of the evaluation in a way that may be more appropriate for epidemiological forecasts, while preserving the propriety of the score. Many different transformations may be appropriate and useful, depending on the exact context, the desired focus of the evaluation, and specific aspects of the forecasts that forecast consumers care most about (see a broader discussion in Section \ref{sec:discussion}).

For conceptual clarity and to allow for a more in-depth discussion, we focus mostly on the natural logarithm as a particular transformation (referred to as the log-transformtation in the remainder of this manuscript) in the context of epidemic phenomena. The CRPS after applying a log-transformation can be computed as follows:
%
\begin{linenomath*}
\begin{equation*}
    \text{CRPS}(F_{\log}, \log y) = \int_{-\infty}^\infty \left( F_{\log}(x) - 1(x \geq \log y) \right)^2 dx,
\end{equation*}    
\end{linenomath*}
%
where $F_{\log}$ is the log-transformed predictive distribution. Instead of a score representing the magnitude of absolute errors, applying a log-transformation prior to the CRPS yields a score which a) measures relative error (see Section \ref{sec:methods:relative}), b) provides a measure for how well a forecast captures the exponential growth rate of the target quantity (see Section \ref{sec:methods:growthrate}) and c) is less dependent on the expected order of magnitude of the quantity to be predicted (see Section \ref{sec:methods:vst}). 
We therefore argue that such evaluations on the logarithmic scale should complement the prevailing evaluations on the natural scale. 
Other transformations may likewise be of interest. We briefly explore the square root transformation as an alternative transformation. 
Our analysis mostly focuses on the CRPS (or WIS) as an evaluation metric for probabilistic forecasts, given its widespread use throughout the COVID-19 pandemic.

The remainder of the article is structured as follows. In Sections \ref{sec:methods:relative}--\ref{sec:methods:vst} we provide some mathematical intuition on applying the log-transformation prior to evaluating the CRPS, highlighting the connections to relative error measures, the epidemic growth rate and variance stabilizing transformations.
We then discuss practical considerations for applying transformations in general and the log-transformation in particular (Section \ref{sec:methods:considerations}) and the effect of the log-transformation on forecast rankings (Section \ref{sec:methods:rankings}). To analyse the real-world implications of the log-transformation we use forecasts submitted to the European COVID-19 Forecast Hub  \citep[ Section \ref{sec:HUB}]{europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022}. Finally, we provide scoring recommendations, discuss alternative transformations that may be useful in different contexts, and suggest further research avenues (Section \ref{sec:discussion}).

@@

To illustrate the effect of applying the natural logarithm prior to evaluating forecasts we consider the absolute error, which the CRPS and WIS generalize to probabilistic forecasts. We assume strictly positive support (meaning that no specific handling of zero values is needed), a restriction we will address when applying this transformation in practice. When considering a point forecast $\hat{y}$ for a quantity of interest $y$, such that 
%
\begin{linenomath*}
\begin{equation*}
y = \hat{y} + \varepsilon,
\end{equation*}
\end{linenomath*}
the absolute error is given by $|\varepsilon|$. When taking the logarithm of the forecast and the observation first, thus considering 
\begin{linenomath*}
\begin{equation*}
\log y = \log \hat{y} + \varepsilon^*,
\end{equation*}
\end{linenomath*}
the resulting absolute error $\left|\varepsilon^*\right|$ can be interpreted as an approximation of various common relative error measures. Using that $\log(a) \approx a - 1$ if $a$ is close to 1, we get
$$
|\varepsilon^*| = |\log \hat{y} - \log y| = \left|\log\left(\frac{\hat{y}}{y}\right) \right| \ \ \stackrel{\text{if } \hat{y} \ \approx \ y}{\approx} \ \ \left| \frac{\hat{y}}{y} - 1 \right| \ \ = \ \ \left| \frac{\hat{y} - y}{y} \right|.
$$
The absolute error after log transforming is thus an approximation of the \textit{absolute percentage error} \citep[APE,][]{gneitingMakingEvaluatingPoint2011a} as long as forecast and observation are close. As we assumed that $\hat{y} \approx y$, we can also interpret it as an approximation of the \textit{relative error} (RE)
$$
\left| \frac{\hat{y} - y}{\hat{y}} \right|
$$
and the \textit{symmetric absolute percentage error} (SAPE)
$$
\left| \ \frac{\hat{y} - y}{y/2 + \hat{y}/2} \ \right|.
$$
As Figure \ref{fig:SAPE} shows, the alignment with the SAPE is in fact the closest and holds quite well even if predicted and observed value differ by a factor of two or three. Generalising to probabilistic forecasts, the CRPS applied to log-transformed forecasts and outcomes can thus be seen as a probabilistic counterpart to the symmetric absolute percentage error, which offers an appealing intuitive interpretation.

@@

Another interpretation for the log-transform is possible if the generative process is framed as exponential with a time-varying growth rate $r(t)$~\citep[see, e.g.,][]{wallingaHowGenerationIntervals2007}, i.e.
\begin{linenomath*}
\begin{equation*}
\frac{d}{dt}y(t) = r(t)y(t)
\end{equation*}
\end{linenomath*}
%
which is solved by
%
\begin{linenomath*}
\begin{equation*}
y(t) = y_0 \exp \left( \int_0^t r(t') dt' \right) = y_0 \exp (\bar{r}t)
\end{equation*}
\end{linenomath*}
where $y_0$ is an initial data point and $\bar{r}$ is the mean of the growth rate between the initial time point $0$ and time $t$.

If a forecast $\hat{y}(t)$ for the value of the time series at time $t$ is issued at time 0 based on the data point $y_0$ then the absolute error after log transformation is
%
\begin{linenomath*}
\begin{align*}
\epsilon^* &= \left| \log \left[ \hat{y}( t ) \right] - \log \left[ y ( t ) \right] \right|\\
           &= \left| \log \left[ y_0 \exp (\bar{\hat{r}} t ) \right] - \log \left[ y_0 \exp (\bar{r}t) \right] \right|\\
           &= t \left| \bar{\hat{r}} - \bar{r} \right|
\end{align*}
\end{linenomath*}
%
where $\bar{\hat{r}}$ is the true mean growth rate and $\bar{r}$ is the forecast mean growth rate. We thus evaluate the error in the mean exponential growth rate, scaled by the length of the time period considered. Again generalising this to the CRPS and WIS implies a probabilistic evaluation of forecasts of the epidemic growth rate.
